Fixed-Shape Ellipse by Three Points

### 2019-10-19 ONPCE/slides.tex
wip

	If the radius is $b$, the angle is a solution:

@@

	\begin{equation*}\label{eq:trpnts}
	\varphi(p, \theta)=\left[\begin{array}{cc}
	\frac{b}{a}&0\\
	0&1
	\end{array}\right]
	\left[\begin{array}{cc}
	\cos{\theta}&\sin{\theta}\\
	-\sin{\theta}&\cos{\theta}
	\end{array}\right]\left[\begin{array}{c}
	p_x\\
	p_y
	\end{array}\right],
	\end{equation*}
	with $p=u, v, w$.

@@

\begin{frame}{Fixed-shape ellipse by three points}
	There is a known formula for the radius of a circumscribed circle:

@@

\begin{frame}{Fixed-shape ellipse by three points}
	There is no clear pattern in $\xi$.
\begin{figure}
	\centering

@@

@@

	\begin{block}{Chebyshev Polynomial}
		$T_n: [-1, 1] \mapsto [-1, 1]$ is the $n$-degree Chebyshev polynomial:
		\begin{equation*}
		T_n(\cos t) = \cos(nt)
		\end{equation*}
	\end{block}
	Also, it can be defined recursively:	
	\begin{align*}
	&T_0(x) = 1\\
	&T_1(x) = x\\
	&T_n(x) = 2xT_{n-1}(x) - T_{n-2}(x)
	\end{align*}

@@

@@

	\begin{itemize}
		\item A simple change of coordinates lets the interpolation to be done on any closed interval!
		\item This can be done in $\bigO(n^2)$.
	\end{itemize}
\end{frame}

@@

	\begin{itemize}
		\item Numerically stable! Way better then polynomials in the power format.

		\item No Runge's Phenomenon, the interpolation converges to $f$.
		\begin{itemize}
			\item $\bigO(n^{-m})$ if $f$ is $m$ times differentiable.
			\item $\bigO(C^n)$, for $C<1$, if $f$ is analytical in a neighborhood of $[-1, 1]$.
		\end{itemize}

@@

	\begin{itemize}
		\item There is no guaranteed way to choose it.
		\item A good rule is to examine the last coefficient.
		\item For a predefined $\epsilon$, choose $n$, such that:

@@

		\item There are other ways like checking the error on a Lobatto grid.
	\end{itemize}
\end{frame}

@@

\begin{frame}{Chebyshev Interpolation}{Roots}
	The roots of a Chebyshev polynomial can be found though determining the eigenvalues of a Chebyshev companion matrix. This is it for $n=5$:

@@

	\begin{itemize}
		\item This matrix is a Hessenberg matrix.
		\item Its eigenvalues can be found by a QR algorithm in $\bigO(n^3)$.
	\end{itemize}
\end{frame}

@@

\begin{frame}{Example}
	An example with $4$ solutions.\cite{andreta}.
	\begin{figure}
		\centering

@@

>>>>

	If the radius is $b$, the angle of rotation is a solution:

@@

	\begin{equation*}\label{eq:trpnts}
	\varphi(p, \theta)=\left[\begin{array}{cc}
	\frac{b}{a}&0\\
	0&1
	\end{array}\right]
	\left[\begin{array}{cc}
	\cos{\theta}&\sin{\theta}\\
	-\sin{\theta}&\cos{\theta}
	\end{array}\right]\left[\begin{array}{c}
	p_x\\
	p_y
	\end{array}\right],
	\end{equation*}
	for $p=u, v, w$.

@@

\begin{frame}{Fixed-shape ellipse by three points}
	There is a known formula for the radius of a circumscribed circle \cite[p.~189]{johnson1960}:

@@

\begin{frame}{Fixed-shape ellipse by three points}
	There is no clear pattern in $\xi$.
	\begin{itemize}
		\item It can be written as $\{\cos^j{x}\sin^k{x} : j, k \in \mathbb{N}\}$.
		\item Degree $6$, at most $12$ roots in $[0, 2\pi)$ \cite[p.~150]{powell}
	\end{itemize}
\begin{figure}
	\centering

@@

\begin{frame}{Polynomial Interpolation}
	It is a way to approximate a function by a simpler one (a polynomial).

	\begin{itemize}
		\item A degree $n$ of the interpolation is determined.
		\item $n+1$ points are chosen, such that the polynomial has to pass through.
		\item Can be calculated using Lagrange's formula.

		\item We can find every root of a polynomial by determining the eigenvalues of a matrix called The Companion Matrix \cite[p.~195]{horn}.

		\item Depending on the points, the interpolation can be a bad approximation. It can get worse even if $n$ is increased (Runge's Phenomenon) \cite[p.~37]{powell}.
	\end{itemize}
\end{frame}

@@

	\begin{block}{Chebyshev Polynomial}
		$T_n: [-1, 1] \mapsto [-1, 1]$ is the $n$-degree Chebyshev polynomial \cite{chebbook}:
		\begin{equation*}
		T_n(\cos t) = \cos(nt)
		\end{equation*}
	\end{block}
	Also, it can be defined recursively:	
	\begin{align*}
	&T_0(x) = 1\\
	&T_1(x) = x\\
	&T_n(x) = 2xT_{n-1}(x) - T_{n-2}(x)
	\end{align*}

@@

\begin{frame}{Chebyshev Interpolation}
\begin{figure}
	\centering

	%\caption{Transforming an ellipse into a circle. T1, T2, and T3 represent the steps of the transformation.}
	\includegraphics[scale=.6]{chebpoly}
	\caption{Chebyshev Polynomials of degree $1, 2, 3, 4$.}
	%	\label{fig:circumscribed-circle}
\end{figure}
\end{frame}

@@

	\begin{itemize}
		\item A simple change of coordinates lets the interpolation to be done on any closed interval!
		\item This can be done in $\bigO(n^3)$ \cite{boyd:2013}.
	\end{itemize}
\end{frame}

@@

	\begin{itemize}
		\item Numerically stable! Way better then polynomials in the power format \cite{gautschi:1979}.

		\item No Runge's Phenomenon, the interpolation converges to $f$.
		\begin{itemize}
			\item $\bigO(n^{-m})$ if $f$ is $m$ times differentiable \cite[p.~28]{gottlieb}.
			\item $\bigO(C^n)$, for $C<1$, if $f$ is analytical in a neighborhood of $[-1, 1]$ \cite{battles:2004}.
		\end{itemize}

@@

	\begin{itemize}
		\item There is no guaranteed way to choose it.
		\item A good rule is to examine the last coefficient (the last coefficient rule-of-thumb \cite[p~.50]{boyd:2001}).
		\item For a predefined $\epsilon$, choose $n$, such that:

@@

		\item There are other ways like checking the error on a Lobatto grid \cite{boyd:2007}.
	\end{itemize}
\end{frame}

@@

\begin{frame}{Chebyshev Interpolation}{Roots}
	The roots of a Chebyshev polynomial can be found though determining the eigenvalues of a Chebyshev companion matrix \cite{boyd:2013}. This is it for $n=5$:

@@

	\begin{itemize}
		\item This matrix is a Hessenberg matrix.
		\item Its eigenvalues can be found by a QR decomposition in $\bigO(n^3)$.
	\end{itemize}
\end{frame}

@@

\begin{frame}{Example}
	An example with $4$ solutions.
	\begin{figure}
		\centering

@@

\begin{frame}[allowframebreaks]
	\frametitle{References}
	\bibliographystyle{amsalpha}
	\bibliography{references}
\end{frame}
